On the entropy of isothermal dusty gas: application to protoplanetary disks

### 2016-08-05 paper.tex
more discussions
starting intro

%\documentclass[12pt]{aastex}
\documentclass[iop, numberedappendix]{emulateapj}
%\documentclass{article}
%\usepackage{draftwatermark}
%\usepackage{paralist}
\shorttitle{untitled}
\shortauthors{M-K.\ Lin}

@@

\section{Introduction}

\input{model}
\input{linear}
\input{result}
%\input{linear_adia}
%\input{numerical}
%\input{application} 
\input{summary}

>>>>

%\documentclass[12pt]{aastex}
\documentclass[iop, numberedappendix]{emulateapj}
%\documentclass{article}
%\usepackage{draftwatermark}
%\usepackage{paralist}
\shorttitle{Isothermal dusty gas dynamics}
\shortauthors{M-K.\ Lin, A.N. Youdin}

@@

\section{Introduction}
%si, sgi as examples
Protoplanetary disks are comprised of a mixture of gas and dust
\citep{chiang10}. Although the overall dust content is typically small
($\sim 1\%$ by mass), their presence can have profound impacts on the
gas dynamics and vice versa.  
%phenomena that depends on gas drag 
%planet formation in dusty disks 
%dust component is building block of planets
%understanding dusty gas dynamics 
%small amount of dust can have significant consequences

%particles 
Current state-of-the-art models of dusty protoplanetary disks
directly simulate pure gas dynamics with explicit Lagragian dust
particles.    
%refs

A simpler framework is to model the dust population as a pressureless
fluid. The advantage is that standard hydrodynamics codes may be
modified by adding a 
%The two-fluid model is also suitable for analytic studies.

%how to model dusty gas?

%two fluid
%one fluid

We draw the physical and mathematical parallel between dusty gas and
adiabatic pure gas. 
%suitable for 
%analytic studies

\input{model}
\input{linear}
\input{result}
\input{summary}

### 2016-08-05 summary.tex
starting intro

@@

\subsubsection{Gravitational instability} %vertical structure 
The addition of dust enhances gravitational
instability (GI). This is because dust particles contribute to the
total disk mass but not thermal pressure, which effectively lowers the
disk temperature \citep[][]{thompson88}. For typical dust-loading
$\epsilon, Z\ll1$, this effect is unimportant. However, if $\epsilon$ is
large (e.g. due to dust settling) then the effective temperature may 
be lowered to enable instability.

As noted in \S\ref{loc_iso_eos}, dust settling causes the effective
temperature to \emph{increase} away from the midplane. This contrasts
to previous studies of GI in vertically stratified disks
\citep[e.g.][]{mamat10, kim12,lin14c} where the temperature decreases
from the midplane. While we expect only the total surface density and
characteristic temperatures are relevant to stability 
\citep{toomre64}, having a non-trivial vertical temperature
structure, induced by dust, may modify the vertical structure of 3D
waves and unstable modes. We will examine this problem in a following
study.

>>>>

\subsection{Implications for numerical simulations}
The dusty/adiabatic gas equivalence means that pure
gas dynamics codes can be applied to simulate locally
isothermal/polytropic dusty gas with minimal modification. In fact, no
modification is needed  
for strictly isothermal or polytropic gas ($c_s$ or $K$ being
constant, respectively) perfectly-coupled to dust. The usual 
adiabatic energy equation is effectively a dust continuity
equation. When $c_s$ or $K$ is not constant and/or $\tstop\neq0$ one
can simply add corresponding source terms in energy equation. In the
large but finite coupling case, this source term is analogous to
radiative diffusion \citep{price15}, which is also standard in many
hydrodynamics codes.

@@

\subsubsection{Gravitational instability} %vertical structure 
The addition of dust enhances gravitational
instability (GI). This is because dust particles contribute to the
total disk mass but not thermal pressure, which effectively lowers the
disk temperature \citep[][]{thompson88,shi13}. For typical dust-loading 
$\epsilon, Z\ll1$, this effect is unimportant. However, if $\epsilon$ is
large (e.g. due to dust settling) then the effective temperature may 
be lowered to enable instability.

As noted in \S\ref{loc_iso_eos}, dust settling causes the effective
temperature to \emph{increase} away from the midplane. This contrasts
to previous studies of GI in vertically stratified disks
\citep[e.g.][]{mamat10, kim12,lin14c} where the temperature decreases
from the midplane. While we expect only the total surface density and
characteristic temperatures are relevant to stability  
\citep{toomre64}, having a non-trivial vertical temperature
structure, induced by dust, may modify the vertical structure of 3D
waves and unstable modes. We will examine this linear problem in a following 
study.
